% Copernicus manuscript
\documentclass[tc, ms, noauthor]{copernicus}

% Copernicus final print
%\documentclass[tc, noauthor]{copernicus}

% Copernicus discussion paper
%\documentclass[tcd, hvmath]{copernicus_discussions}

% Extra packages
\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{framed}

% Recognize utf-8 special characters
\usepackage[utf8]{inputenc}

% All floats at the end of the manuscripts
%\usepackage{endfloat}

% Add to-do notes and enlarge right margin
\usepackage{todonotes}
\usepackage[left=54pt, top=81pt, textwidth=398pt, marginparsep=11pt, marginparwidth=93pt, textheight=625pt, footskip=30pt]{geometry}
\newcommand{\needref}[0]{\todo[color=cyan]{Reference needed}}
\newcommand{\arjen}[2][]{\todo[color=lime,{#1}]{\textbf{Arjen:} #2}}
\newcommand{\constantine}[2][]{\todo[color=yellow,{#1}]{\textbf{Constantine:} #2}}
\newcommand{\qiong}[2][]{\todo[color=pink,{#1}]{\textbf{Qiong:} #2}}
\newcommand{\ask}[2][]{\todo[color=lightgray,{#1}]{\textbf{Julien:} #2}}

% Personal shortcuts
\newcommand{\degC}[0]{\,\unit{\degree C} }
\newcommand{\masl}[0]{\,\unit{m a.s.l.} }

% Set figures repository
\graphicspath{{figures/}}

% ----------------------------------------------------------------------
\begin{document}

\title{Choosing climate forcing for numerical simulations of the Cordilleran ice sheet}
\runningtitle{Cordilleran ice sheet climate forcing}

%\author{Julien Seguinot et al.}
%\affil{Department of Physical Geography and Quaternary Geology, Stockholm University, S-106 91 Stockholm, Sweden}
\runningauthor{J. Seguinot et al.}
\correspondence{J. Seguinot\\ (julien.seguinot@natgeo.su.se)}

\received{}
\pubdiscuss{}
\revised{}
\accepted{}
\published{}

\maketitle
\ask[inline]{\emph{Should the title be changed?}

	\emph{I am open for discussion. As I see it, this paper has two major lines of thought, maybe not so easily combined:
	\begin{enumerate}
		\item Choosing climate forcing for palæo-ice sheet modelling. This is the main point of the paper if only. I am choosing climate forcing for the rest of my PhD as far as I am concern. Choosing climate forcing for other palæo-ice sheet modelling exercises as the reader is (potentially) concerned.
		\item Numerically modelling the Cordilleran ice sheet. This is a point in itself as it has not quite being done so far as far as I am aware of.
	\end{enumerate}
	So I tried my best to include both ideas in the title. Yet this is preliminary...}}

\arjen[inline]{The effect of climate forcing on numerical simulations of the Cordilleran ice sheet at the last glacial maximum}

% ----------------------------------------------------------------------
\begin{abstract}
\todo{This abstract is preliminary. I will rewrite it when I am done with the rest}
We present an ensemble of numerical simulations of the former Cordilleran ice sheet forced by present-day climatologies from five different datasets. Surface air temperature and precipitation rates from the NCEP/NCAR reanalysis, the Climate Forecast System Reanalysis, the ERA-Interim reanalysis and the North American Regional Reanalysis are used to force the ice sheet model PISM. Glacial conditions are mimicked by applying constant temperature offsets over the entire modelling domain, and a positive degree-day model is used to compute surface mass balance. The modelled ice sheet appears highly sensitive to the choice of climate forcing. Best results were obtained from the North American Regional Reanalysis, a high-resolution regional reanalysis which assimilates precipitation.
\end{abstract}
% ----------------------------------------------------------------------

% ----------------------------------------------------------------------
\input{sections/intro}
\input{sections/climate}
\input{sections/model}
\input{sections/results}
\input{sections/discuss}
\input{sections/technote}
% ----------------------------------------------------------------------

% ----------------------------------------------------------------------
\conclusions  %% \conclusions[modified heading if necessary]
\label{sec:conclusion}
\emph{Coming soon...}
% ----------------------------------------------------------------------

% ----------------------------------------------------------------------
%\begin{acknowledgements}
%\end{acknowledgements}
% ----------------------------------------------------------------------

\bibliographystyle{copernicus}
\bibliography{references}
\end{document}

