% response-stumpf.tex
% ----------------------------------------------------------------------
\input{response-header}

\begin{document}
\textbf{Authors' response to A.~Stumpf}
\bigskip

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\newcommand{\sechead}[1]{\bigskip\noindent\textbf{#1}}
\newcommand{\referee}[1]{\bigskip\textcolor{journalname}{\textit{#1}}}
\newcommand{\msquote}[1]{\begin{quote}\textit{#1}\end{quote}}

To A.~Stumpf,

Thank you very much for taking the time to read and comment on our manuscript.
We are very thankful to receive feedback from the glacial geology community,
more particularly so here on The Cryosphere Discussion which traditionally
has a stronger presence in quantitative glaciology.

\referee{%
    Seguinot et al. present a well designed numerical model for the Cordilleran
    Ice Sheet (CIS) in North America for the last two glaciations occurring
    over the past 120000 years. Although the constraints on such a model are
    not yet all fully understood, their proposed simulations attempt to take
    account all the complexity of the glacial system, and utilizes a variety of
    data sets. The modeling confirms what geologists have observed in the
    field; there was a fast decay of the CIS during MIS~2 and MIS~4 and
    non-glacial conditions existed during MIS (Olympia Nonglacial Interval;
    \citealp[e.g.,][]{Plouffe.Jette.1997}). I applaud them for undertaking of
    such a difficult task, and notifying the reader where input data is sparse
    or inconclusive.}

Thank you very much for this positive summary of our work and for your
appreciative words!

\referee{%
    I provide the following general comments and observations which may help
    the authors in revising the manuscript for final publication. Many of these
    points are both my personal suggestions and also the recommendations of
    others currently researching the CIS or who have completed studies in the
    past.}

\referee{%
    1) With the large amount of research that has been undertaken to map the
    landforms and deposits of the CIS by federal and provincial scientists,
    academic faculty, and undergraduate and graduate students and determine the
    extent, volume and dynamics ice sheet, the impression left on the reader by
    the opening sentence would be incorrect. These studies have greatly
    advanced our understanding of the CIS, and could be an important dataset to
    test against the modeling. \citet{Stumpf.etal.2014} provides a list of some
    of these studies.}

By no means our opening sentence was meant to undervalue the large amount of
geological work performed on the Cordilleran ice sheet, much of which has been
reviewed and discussed against our model results in later parts of the
manuscript. Despite all the work done, our impression is that the Cordilleran
ice sheet, due to its complexity, remains less understood than its Laurentide
and Eurasian counterparts, especially when it comes to reconstructing the
dynamics of the different phases of advance and retreat of the ice sheet
through the last glacial cycle. This is the impression that we tried to convey
in the first sentence of the abstract. To avoid any further misunderstanding,
we replaced the opening two sentences with:

\msquote{%
    After more than a century of geological research, the Cordilleran ice sheet
    of North America remains the least understood among Pleistocene ice sheets
    in terms of its former extent, volume and dynamics. Because of the
    irregular topography on which the ice sheet formed, geological studies have
    often had only local or regional relevance, and evidenced such a complexity
    that whole ice-sheet reconstructions of advance and retreat patterns are
    lacking.}

\referee{%
    2) To help the reader better understand the maps presented, I would
    recommend some spatial information be added (e.g., latitude/longitude
    grids; political boundaries; lakes and rivers, place names etc..).}

\todo{%
    Add info where it will not disturb the figures. Or perhaps a larger
    location map like in the Kappa?}

\referee{%
    3) Was the model tested against regional-scale ground-based data
    \citep[e.g.,][]{Ferbey.etal.2013} to constrain ice divide positions, ice
    flow direction, and ice sheet thickness?}

\todo{%
    Nice dataset. Not quantitatively tested, if that is what the comment means.
    In my understanding, there is very little reliable info about ice sheet
    thickness, ice divide positions are reconstructions that always involve
    some level of interpretation. Comparing flow direction against
    \citet{Ferbey.etal.2013} would be very nice but I don't think we are this
    confident in the basal velocities for such a comparison yet.}

\referee{%
    4) For the central sector of the CIS, \citet{Stumpf.etal.2000} provides
    some insight into the chronology and effectiveness of glacial erosional
    during the MIS. In figure~5, and in the accompanying text, they describe
    how landforms on the surface formed. For example, in lake valleys east of
    the Skeena Mountains, it appeared the major glacial streamlined landforms
    were formed during a longer glacial advance phase, with ice flow
    paralleling the valleys, and later flows, some perpendicular the valley
    flow, only weakly impacted them.}

\todo{%
    Add reference to \citet{Stumpf.etal.2000}...}

\referee{%
    5) \citet{Stumpf.etal.2000} was the first study to extensively document a
    predominant westerly directed ice flow across high elevations in the Skeena
    and Coast Mountains. This flow direction appeared to continue into the
    late-glacial period. Other subsequent studies also confirm that
    late-glacial readvance eastward out of these mountains and retreat of ice
    margins westward into these mountains was limited.}

\todo{%
    Add reference to \citet{Stumpf.etal.2000}...}

% ----------------------------------------------------------------------

\begin{thebibliography}{69}

\bibitem[{Ferbey et~al.(2013)}]{Ferbey.etal.2013}
Ferbey, T., Arnold, H., and Hickin, A.: Ice-flow indicator compilation, British
  Columbia., British Columbia Geol. Surv., Victoria, BC, Open-File 2013-06,
  \url{http://www.em.gov.bc.ca/Mining/Geoscience/PublicationsCatalogue/OpenFiles/2013/Pages/2013-06.aspx},
  2013.

\bibitem[{Seguinot et~al.(2014)}]{Seguinot.etal.2014}
Seguinot, J., Khroulev, C., Rogozhina, I., Stroeven, A.~P., and Zhang, Q.: The
  effect of climate forcing on numerical simulations of the {C}ordilleran ice
  sheet at the {L}ast {G}lacial {M}aximum, The Cryosphere, 8, 1087--1103,
  \doi{10.5194/tc-8-1087-2014}, 2014.

\bibitem[{Stumpf et~al.(2000)}]{Stumpf.etal.2000}
Stumpf, A.~J., Broster, B.~E., and Levson, V.~M.: Multiphase flow of the late
  Wisconsinan Cordilleran ice sheet in western Canada, Geol. Soc. Am. Bull.,
  112, 1850--1863, \doi{10.1130/0016-7606(2000)112<1850:mfotlw>2.0.co;2}, 2000.

\bibitem[{Stumpf et~al.(2014)}]{Stumpf.etal.2014}
Stumpf, A.~J., Ferbey, T., Plouffe, A., Clague, J.~J., Ward, B.~C., Paulen,
  R.~C., and Bush, A.~B.: Discussion: ``Streamlined erosional residuals and
  drumlins in central British Columbia, Canada'' by J. Donald {McClenagan},
  (2013) Geomorphology 189, 41--54, Geomorphology, 209, 147--150,
  \doi{10.1016/j.geomorph.2013.10.019}, 2014.

\bibitem[{Plouffe and Jett{\'{e}}(1997)}]{Plouffe.Jette.1997}
Plouffe, A. and Jett{\'{e}}, H.: Middle Wisconsinan sediments and paleoecology
  of central British Columbia: sites at Necoslie and Nautley rivers, Can. J.
  Earth Sci., 34, 200--208, \doi{10.1139/e17-016}, 1997.

\end{thebibliography}

% ----------------------------------------------------------------------
% Interactive comment text ends
% ----------------------------------------------------------------------

\end{document}
